---
title: "Trace the datascape"
subtitle: "Descriptive assessment of datasets"
date: "2024-02-09"
categories: [ggplot2, skimr, janitor, knitr, summarize(), group_by(), count(), tabyl(), kable(), geom_*()]
webr:
  show-startup-message: false
  packages: ['dplyr', 'knitr', 'ggplot2', 'skimr', 'corpora', 'janitor', 'kableExtra']
  message: false
---

## Overview

- Statistical overviews
- Data summaries
- Visual summaries
- Lab 03: Trace the datascape

```{r}
#| label: slides-setup
#| include: false

# Load packages
library(dplyr)
library(knitr)
library(ggplot2)
library(skimr)
library(corpora)
library(janitor)
library(kableExtra)

# Read the dataset from the `corpora` package
brown_fam_df <-
  corpora::PassiveBrownFam |> # reference the dataset
  as_tibble() # convert to a tibble

# Rename variables
brown_fam_df <-
  brown_fam_df |> # pass the original dataset
  rename( # rename variables: new_name = old_name
    variety = lang,
    num_words = n.words,
    act_verbs = act,
    pass_verbs = pass,
    verbs = verbs,
    pct_passive = p.pass
  ) |>
  select(-corpus, -section, -period, -id) |>
  mutate(id = row_number(), genre = stringr::str_to_title(genre)) |>
  select(id, everything())
```

```{webr-r}
#| context: setup

# Read the dataset from the `corpora` package
brown_fam_df <-
  corpora::PassiveBrownFam |> # reference the dataset
  as_tibble() # convert to a tibble

# Rename variables
brown_fam_df <-
  brown_fam_df |> # pass the original dataset
  rename( # rename variables: new_name = old_name
    variety = lang,
    num_words = n.words,
    act_verbs = act,
    pass_verbs = pass,
    verbs = verbs,
    pct_passive = p.pass
  ) |>
  select(-corpus, -section, -period, -id) |>
  mutate(id = row_number(), genre = stringr::str_to_title(genre)) |>
  select(id, everything())
```

## Statistical overviews{.smaller .scrollable}

::: {.panel-tabset}
### skim()
```{webr-r}
# Use skim() to get all variables
brown_fam_df |>
  skim()

# Use skim() |> yank("var_type") to get specific types (e.g. "character", "numeric", etc.)
brown_fam_df |>
  skim() |>
  yank("numeric")
```

### summarize()
```{webr-r}
# Use summarize() to get summary statistics
brown_fam_df |>
  summarize(
    mean = mean(num_words),
    sd = sd(num_words),
    median = median(num_words),
    iqr = IQR(num_words)
  )
```
:::

## Data summaries{.smaller .scrollable}

::: {.panel-tabset}
### count()
```{webr-r}
# Use count() to get frequency tables
brown_fam_df |>
  count(genre, variety)
```

### tabyl()
```{webr-r}
# Use tabyl() to get contingency tables
brown_fam_df |>
  tabyl(genre, variety)
```

### group_by() + summarize()
```{webr-r}
# Use group_by() + summarize() to get pivot tables
brown_fam_df |>
  group_by(genre) |>
  summarize(
    mean_words = mean(num_words),
    sd_words = sd(num_words)
  )
```
:::

## Data summaries to tables{.smaller .scrollable}

::: {.panel-tabset}
### kable()
```{webr-r}
variety_word_stats <-
  brown_fam_df |>
  group_by(variety) |>
  summarize(
    mean_words = mean(num_words),
    sd_words = sd(num_words)
  )

variety_word_stats |>
  kable()
```

### kable() results
```{r}
#| label: kable-data
#| echo: false

variety_word_stats <-
  brown_fam_df |>
  group_by(variety) |>
  summarize(
    mean_words = mean(num_words),
    sd_words = sd(num_words)
  )
```

```{{r}}
#| label: tbl-kable-results
#| tbl-cap: "Variety word statistics"

variety_word_stats |>
  kable()
```

```{r}
#| label: tbl-kable-results
#| tbl-cap: "Variety word statistics"
#| echo: false

variety_word_stats |>
  kable()
```
:::

## Visual summaries{.smaller .scrollable}

Density plots

::: {.panel-tabset}
### One variable
```{webr-r}
# Use geom_density() to visualize the distribution of num_words
brown_fam_df |>
  ggplot(aes(x = num_words)) +   # specify the x-axis variable
  geom_density()
```

### Two variables{.smaller .scrollable}
```{webr-r}
# Use geom_density() to visualize the distribution of num_words by variety
brown_fam_df |>
  ggplot(aes(x = num_words, fill = variety)) + # specify the x-axis and fill variables
  geom_density(alpha = 0.5)                   # specify alpha for transparency
```
:::

## Visual summaries{.smaller .scrollable}

Boxplots

::: {.panel-tabset}
### One variable
```{webr-r}
# Use geom_boxplot() to visualize the distribution of pass_verbs
brown_fam_df |>
  ggplot(aes(x = pass_verbs)) + # specify the x-axis
  geom_boxplot()                # create the boxplot
```

### Two variables
```{webr-r}
# Use geom_boxplot() to visualize the distribution of pass_verbs by variety
brown_fam_df |>
  ggplot(aes(x = variety, y = pass_verbs)) + # specify the x-axis and y-axis
  geom_boxplot()                            # create the boxplot
```
:::

## Visual summaries{.smaller .scrollable}

Scatterplots

::: {.panel-tabset}
### Two variables
```{webr-r}
# Use geom_point() to visualize the relationship between num_words and verbs
brown_fam_df |>
  ggplot(aes(x = num_words, y = verbs)) + # specify the x-axis and y-axis
  geom_point()                                # create the scatterplot
```
### Two variables with trend line
```{webr-r}
# Use geom_point() and geom_smooth() to visualize the relationship between num_words and verbs
brown_fam_df |>
  ggplot(aes(x = num_words, y = verbs)) + # specify the x-axis and y-axis
  geom_point() +                              # create the scatterplot
  geom_smooth(method = "lm")                  # add a trend line
```
:::

# Lab 03: Trace the datascape

## Setup

- Clone the repository to RStudio as a new project
- Follow the instructions in the `README.md` file
